\PYGEOMETRY{}: a Python library for the creation of Monte Carlo radiation transport physical geometries

### 2020-09-15 public/paper/pyg4ometry.tex
Paper update

STEP and IGES files can be loaded into \PYGEOMETRY{}, via an interface 
based on FreeCAD. FreeCAD is an open source CAD/CAM program, which 
in turn is based on OpenCASCADE. FreeCAD allows for scripting in Python 
and acts as a simple to use interface to OpenCASCADE.  A STEP CAD file 
could be considered as a hierarchical tree of part features and part assemblies, 
where a part assembly is a collection of part features. A part feature can be used 
to create a triangular mesh which can be used to create a \PYGEOMETRY{} tessellated solid. The placement 
of the part feature is extracted from the STEP file and used to create an 
appropriate physical volume. Assignment of materials and visualisation 
attributes must be performed by  the user after conversion to GDML as rarely 
will CAD/CAM packages included the detailed information required for MCRT 
codes. Listing~\ref{lst:pythonCADLoading} shows 
how \PYGEOMETRY{} can load step files.
\begin{lstlisting}[caption={A simple \PYGEOMETRY{} Python script to load a STEP file.},label={lst:pythonCADLoading}, language=Python]
    reader = pyg4ometry.freecad.Reader("CadFileName.step")
    reader.relabelModel()
    reader.convertFlat()
    logical = reader.getRegistry().getWorldVolume()
\end{lstlisting}
Compared to other file readers, two additional steps are required \pyinline{relabelModel} and \pyinline{convertFlat}. CAD model 
part names can contain characters which are not allowed in Python dictionaries so need to be replaced by \pyinline{relabelModel}. 
CAD models might also have a hierarchy of parts and assemblies, these are converted without this structure by \pyinline{convertFlat}.
In general there is no requirement to avoid geometric overlap  of parts in a CAD file. This will result in overlaps of the solids in 
the converted tessellated solids. This is avoided by shrinking each solid, this is done by computing the a normal ${\bf n}$  for each 
vertex ${\bf v}$ and shifting its position by $\epsilon {\bf n}$, so the new vertices are ${\bf v} - \epsilon {\bf n}$. The degree of shrinking 
is user controllable.

@@

There are more complex extensions that can be considered for inclusion into \PYGEOMETRY{}.
The meshes created by \PYGEOMETRY{} are generally of very high quality and can be used for a 
wide range of applications. An idea already being developed is the export of the geometry mesh data to
data formats used in augmented or virtual reality software to create interactive visualisations of MCRT 
simulations.  Triangular meshes also have application  for GPU accelerated photon tracking in 
liquid noble dark matter detectors. Paraview/VTK are being standard software for complex 3D visualisation and 
the ability to write geometry to formats readily loaded and manipulated by these programs will 
significantly aid the presentation of geometry along with the results of the MCRT simulations.

>>>>

STEP and IGES files can be loaded into \PYGEOMETRY{}, via an interface 
based on FreeCAD. FreeCAD is an open source CAD/CAM program, which 
in turn is based on OpenCASCADE. FreeCAD allows for scripting in Python 
and acts as a simple to use interface to OpenCASCADE.  A STEP CAD file 
could be considered as a hierarchical tree of parts and part assemblies, 
where a part assembly is a collection of part features. A part feature can be used 
to create a triangular mesh which can be used to create a \PYGEOMETRY{} tessellated solid. The placement 
of the part feature is extracted from the STEP file and used to create an 
appropriate physical volume. Assignment of materials and visualisation 
attributes must be performed by  the user after conversion to GDML as rarely 
will CAD/CAM packages included the detailed information required for MCRT 
codes. Listing~\ref{lst:pythonCADLoading} shows 
how \PYGEOMETRY{} can load step files.
\begin{lstlisting}[caption={A simple \PYGEOMETRY{} Python script to load a STEP file.},label={lst:pythonCADLoading}, language=Python]
    reader = pyg4ometry.freecad.Reader("CadFileName.step")
    reader.relabelModel()
    reader.convertFlat()
    logical = reader.getRegistry().getWorldVolume()
\end{lstlisting}
Compared to other file readers, two additional steps are required \pyinline{relabelModel} and \pyinline{convertFlat}. CAD model 
part names can contain characters which are not allowed in Python dictionaries so need to be replaced by \pyinline{relabelModel}. 
CAD models might also have a hierarchy of parts and assemblies, these are converted without this structure by \pyinline{convertFlat}.
In general there is no requirement to avoid geometric overlap  of parts in a CAD file. This will result in overlaps of the solids in 
the converted tessellated solids. This is avoided by shrinking each solid, this is done by computing the a normal ${\bf n}$  for each 
vertex ${\bf v}$ and shifting its position by $\epsilon {\bf n}$, so the new vertices are ${\bf v} - \epsilon {\bf n}$. The degree of shrinking 
is user controllable.

@@

There are more complex extensions that can be considered for inclusion into \PYGEOMETRY{}.
The meshes created by \PYGEOMETRY{} are generally of very high quality and can be used for a 
wide range of applications. An idea already being developed is the export of the geometry mesh data to
data formats used in augmented or virtual reality software to create interactive visualisations of MCRT 
simulations.  Triangular meshes also have application  for GPU accelerated photon tracking in 
liquid noble dark matter detectors. Paraview/VTK are becoming standard software for complex 3D visualisation and 
the ability to write geometry to formats readily loaded and manipulated by these programs will 
significantly aid the presentation of geometry along with the results of the MCRT simulations.
